---
title: "CARS_MATH3710"
author: "Cody Frisby"
date: "December 15, 2015"
output: 
  html_document: 
    keep_md: yes
---
Let's get started with R.  This first data set is small so we have the data loaded into two variables: minutes and units.  We then put them together with the cbind function and store the result in a new variable called repair.

```{r}
minutes <- c(23,29,49,64,74,87,96,97,109,119,149,145,154,166, NA)
units <- c(1,2,3,4,4,5,6,6,7,8,9,9,10,10,3)
repair <- as.data.frame(cbind(minutes, units))
head(repair)
summary(repair)
```

Above is a summary of the data table "repair".
Now let's plot the data look at the variable minutes being modeled as a function of units.

```{r}
attach(repair) #attaching the data makes our code below simpler.
plot(units, minutes, bg = "lightblue", 
     col = "black", cex = 1.1, pch = 21,frame = FALSE)
abline(lm(minutes ~ units, data = repair), lwd = 2)
cor(repair, use = "complete.obs")[2]
fit <- lm(minutes ~ units, data = repair)
summary(fit)
hist(residuals(fit), col = "green", freq = FALSE)
```

Below is a plot of the residuals
```{r}
par(mfrow=c(2,2)) #setup your window to fit the next plots
plot(fit)
```
