---
title: "Car MPG"
author: "ccf"
date: "January 17, 2016"
output: 
  html_document: 
    keep_md: yes
---

This is an R markdown file for the car mpg data shown in class.  We will reporduce some of the SAS output using R and Rstudio.

First we need to read the data into R.

```{r}
carmpg <- read.table("~/Documents/MATH3710/cars/MPGData.txt", header = TRUE)
summary(carmpg)
cor(carmpg)
```


Above are simple summary statistics of the data.

Now let's look at a scatter plot of the data.

```{r}
plot(MPG ~ Weight, data=carmpg, bg = "lightblue", 
     col = "black", cex = 1.5, pch = 21)
text(x=1800, y=15, paste0("Cor = ",round(cor(carmpg)[2], 4), cex = 0.7))
```


Now we fit a linear model in R using the lm function. We will store it in a variable for more use.

```{r}
fit <- lm(MPG ~ Weight, data = carmpg)
summary(fit) # displays a summary of the model.
```


Now we will use R to check the assumptions of the model. First we will store the standardized residuals from "fit" and store them in a variable.

```{r}
e <- rstandard(fit)
summary(e)
```

More diagnostic plots of our model:
```{r}
qqnorm(e)
qqline(e)
# Residuals vs. x
x <- carmpg$Weight
plot(x, e, bg = "lightblue", 
     col = "black", cex = 1.1, pch = 21,frame = FALSE)
abline(h = 0, lwd = 2)
```

One more plot.  Here we will look at a histogram of the residuals.

```{r}
hist(e, breaks = 25, col = "green", freq = FALSE, xlim = c(-4,4), main = "Histogram of Standardized Residuals", xlab = "rstandard")
```

