---
title: "Climate Change"
author: "ccf"
date: "January 20, 2016"
output: 
  html_document: 
    keep_md: yes
---

This is an R Markdwon (Rmd) file for the climate change data shared in class.  We will attempt to translate some of the SAS code into R.

```{r}
climate <- read.table("~/Documents/MATH3710/climate/Climate.txt", header = TRUE)
climate$co2_cent <- climate$co2 - mean(climate$co2)
```


Above are simple summary statistics of the data.

Now let's look at a scatter plot of the data.

```{r}
plot(globaltemp ~ co2, data = climate, bg = "lightblue", 
     col = "black", cex = 1.5, pch = 21)
```


Now we fit a linear model in R using the lm function. We will store it in a variable for more use.

```{r}
fit <- lm(globaltemp ~ co2, data = climate)
summary(fit) # displays a summary of the model.
```

Here we will plot the data with confidence intervals too.  This is quite involved in R so the code is quite long.

```{r}
x <- climate$co2
y <- climate$globaltemp
plot(y~x,type="n")
m <- lm(y~x)
wx <- par("usr")[1:2]
new.x <- seq(wx[1],wx[2],len=100)
pred <- predict(m, new=data.frame(x=new.x), interval="conf")
lines(new.x,pred[,"fit"],lwd=2)
lines(new.x,pred[,"lwr"],lty=3)
lines(new.x,pred[,"upr"],lty=3)
points(x,y,pch=16,col="steelblue")
# now to add the prediction intervals
predp <- predict(m, new=data.frame(x=new.x), interval = "pred")
lines(new.x, predp[,"lwr"], lty=2)
lines(new.x, predp[,"upr"], lty=2)
title(main="Fit Plot for Temp")
```


Now we will use R to check the assumptions of the model. First we will store the standardized residuals from "fit" and store them in a variable.

```{r}
e <- rstandard(fit)
summary(e)
# plot residuals by co2
plot(climate$co2, e, bg = "lightblue", col = "black", pch = 21,
     xlab = "Co2", ylab = "Residuals", main = "Residuals vs Co2")
abline(h=0)
```

More diagnostic plots of our model residuals:
```{r}
qqnorm(e)
qqline(e)
hist(e, breaks = 25, col = "green", freq = FALSE, xlim = c(-4,4), 
     main = "Histogram of Standardized Residuals", xlab = "rstandard")
```
